---
title: "EDA_PROJECT2"
author: "BHubler"
date: "October 24, 2015"
output: html_document
---
```{r}
setwd("/Users/berlyn/Google Drive/Coursera R/EDA")
```

PM2.5 Emissions Data (summarySCC_PM25.rds): This file contains a data frame with all of the PM2.5 emissions data for 1999, 2002, 2005, and 2008. For each year, the table contains number of tons of PM2.5 emitted from a specific type of source for the entire year. Here are the first few rows.

```{r}
# read in the data files
NEI <- readRDS("./exdata-data-NEI_data/summarySCC_PM25.rds")
SCC <- readRDS("./exdata-data-NEI_data/Source_Classification_Code.rds")
```

## Assignment

The overall goal of this assignment is to explore the National Emissions Inventory database and see what it say about fine particulate matter pollution in the United states over the 10-year period 1999–2008. You may use any R package you want to support your analysis.

## Questions

You must address the following questions and tasks in your exploratory analysis. For each question/task you will need to make a single plot. Unless specified, you can use any plotting system in R to make your plot.

### 1
Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r, echo=FALSE}
# subset the data for each year
plot1_1999dat <- NEI[(NEI$year == 1999), ]
plot1_2002dat <- NEI[(NEI$year == 2002), ]
plot1_2005dat <- NEI[(NEI$year == 2005), ]
plot1_2008dat <- NEI[(NEI$year == 2008), ]

# make new data sets with years (x) and total emissions(y) for that year
x <- c(1999, 2002, 2005, 2008)
y <- c(sum(plot1_1999dat$Emissions), sum(plot1_2002dat$Emissions), sum(plot1_2005dat$Emissions), sum(plot1_2008dat$Emissions))

# plot data
plot(x,y, type="p", main="Total PM2.5 Emissions - All Sources", xlab="Year", ylab="Total PM2.5 Emissions [tons]")
abline(lm(y ~ x))

# save plot
dev.copy(png, file="plot1.png", height=480, width=480)
dev.off()
```

### 2
Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510") from 1999 to 2008? Use the base plotting system to make a plot answering this question.

```{r, echo=FALSE}
# read data
NEI <- readRDS("./exdata-data-NEI_data/summarySCC_PM25.rds")

# subset the data for each year & fips == "24510"
plot2_1999dat <- NEI[(NEI$year == 1999 & NEI$fips == "24510"), ]
plot2_2002dat <- NEI[(NEI$year == 2002 & NEI$fips == "24510"), ]
plot2_2005dat <- NEI[(NEI$year == 2005 & NEI$fips == "24510"), ]
plot2_2008dat <- NEI[(NEI$year == 2008 & NEI$fips == "24510"), ]

# make new data sets with years (x) and total emissions(y) for that year & region
x <- c(1999, 2002, 2005, 2008)
y <- c(sum(plot2_1999dat$Emissions), sum(plot2_2002dat$Emissions), sum(plot2_2005dat$Emissions), sum(plot2_2008dat$Emissions))

# plot data
plot(x,y, type="p", main="Total PM2.5 Emissions Baltimore, MD - All Sources", xlab="Year", ylab="Total PM2.5 Emissions [tons]")
abline(lm(y ~ x))

# save plot
dev.copy(png, file="plot2.png", height=480, width=480)
dev.off()

```

### 3
Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008? Use the ggplot2 plotting system to make a plot answer this question.

```{r, echo=FALSE}
# read data

#NEI <- readRDS("./exdata-data-NEI_data/summarySCC_PM25.rds")

# subset the data for each source type
# need a better system than this one....

#plot3_1999dat <- NEI[(NEI$year == 1999 & NEI$type == "POINT"), ]


# make a panel plot for each type (4 total)
# 1. x = year, y = total emissions for POINT
# 2. x = year, y = total emissions for NONPOINT
# 3. x = year, y = total emissions for ON-ROAD
# 4. x = year, y = total emissions for NON-ROAD

#x <- c(1999, 2002, 2005, 2008)
#y <- 

# plot data using ggplot2
#library(ggplot2)


# save plot
#dev.copy(png, file="plot3.png", height=480, width=480)
#dev.off()

```


### 4
Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

### 5
How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

### 6
Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?


# Making and Submitting Plots

For each plot you should

Construct the plot and save it to a PNG file.

Create a separate R code file (plot1.R, plot2.R, etc.) that constructs the corresponding plot, i.e. code in plot1.R constructs the plot1.png plot. Your code file should include code for reading the data so that the plot can be fully reproduced. You must also include the code that creates the PNG file. Only include the code for a single plot (i.e. plot1.R should only include code for producing plot1.png)

Upload the PNG file on the Assignment submission page

Copy and paste the R code from the corresponding R file into the text box at the appropriate point in the peer assessment.